### 2016-11-22 ResearchStatement/main.tex
Back on just 2 pages
Language fixes

Cartesian Genetic Programming (CGP) is a method for evolving digital circuits
and other functions that has had continual use for almost two decades. Part of its
success is that unlike other methods it tends to produce terse results. Through
careful analysis, I was able to determine the reason for this desirable outcome
was actually an unintended consequence of a seemingly unrelated design
decision~\cite{goldman:2013:ordering,goldman:2015:cgpanalysis}. From this finding
I was able to tweak that decision to improve performance, maintain
terse solutions, and remove a parameter from user configuration.

The primary focus of my dissertation was the creation of the
Parameter-less Population Pyramid (P3), an optimization algorithm
that requires no problem specific configuration~\cite{goldman:2014:p3,goldman:2015:fastp3,goldman:2016:p3hiff}.
Its design was inspired by my previous analysis of model building optimization algorithms~\cite{goldman:2012:ltga}.
P3 combines model building, local search, and novel population structuring to minimize the amount
of stochastic exploration.
P3 is able to find the global optimum in less time than previous techniques, as well as
maintain a better ``current best'' solution as optimization progresses, across a diverse set of problems.
This quality is on top
of the fact that P3 requires no configuration.
By exploiting additional information on a subset of problems, I was able to further
improve P3's efficiency by two orders of magnitude~\cite{goldman:2015:GBO}.
Working with a collaborator, we were able to develop a generic method using the same principles
that solves many common problems in the optimization literature in $O(N)$ time~\cite{whitley:2016:mkl}.

@@

\section{Future Work}
Any application that requires a large number of expert decisions be made at low
cost can potentially benefit from ML. Unfortunately, our inability to understand
why ML makes the decisions it does can prevent acceptance of its use. This is
especially true when it comes to using ML to help the poorest and most disadvantaged
members of our society.

>>>>

Cartesian Genetic Programming (CGP) is a method for evolving digital circuits
and other functions that has had continual use for almost two decades. Part of CGP's
success is that it tends to produce terse designs. Through
careful analysis, I was able to determine the reason for this desirable outcome
was actually an unintended consequence of a seemingly unrelated design
decision~\cite{goldman:2013:ordering,goldman:2015:cgpanalysis}. From this finding
I was able to tweak that decision to improve performance, maintain
terse solutions, and remove a parameter from user configuration.

The primary focus of my dissertation was the creation of the
Parameter-less Population Pyramid (P3), an optimization algorithm
that requires no problem specific configuration~\cite{goldman:2014:p3,goldman:2015:fastp3,goldman:2016:p3hiff}.
Its design was inspired by my previous analysis of model building optimization algorithms~\cite{goldman:2012:ltga}.
P3 combines model building, local search, and novel population structuring to minimize the amount
of stochastic exploration.
P3 is able to find the global optimum in less time than previous techniques, as well as
maintain a better ``current best'' solution as optimization progresses, across a diverse set of problems.
This quality is on top
of the fact that P3 requires no configuration.
By exploiting additional information on a subset of problems, I was able to further
improve P3's efficiency by two orders of magnitude~\cite{goldman:2015:GBO}.
%Working with a collaborator, we were able to develop a generic method using the same principles
%that solves many common problems in the optimization literature in $O(N)$ time~\cite{whitley:2016:mkl}.

@@

\section{Future Work}
Any application that requires a large number of expert decisions to be made at low
cost can potentially benefit from ML. Unfortunately, our inability to understand
why a learned model makes the decisions it does can prevent acceptance of its use. This is
especially true when it comes to using ML to help the poorest and most disadvantaged
members of our society.
